---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---
```{r intro, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 
knitr::opts_knit$set(root.dir = "C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_Global/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(readxl)))
suppressMessages(library(plotly))
suppressMessages(library(Matrix))
suppressMessages(library(stringr))
suppressMessages(library(stringi))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(zoo)))
```

```{r dtaWHOregion}
#WHO region classification data from Chelsea Taylor 11/18/2020
dtawhoregion<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_Global/WHO Country Regions.xlsx")
str(dtawhoregion)

dtawhoregion<-dtawhoregion%>%
    rename(country = "Country Name")%>%
    select(country, WHO_REGION)

#prep for join/merge
dtawhoregion<-dtawhoregion%>%
    mutate(
        country=ifelse(country=="Democratic People's Republic of Korea","Dem. People's Republic of Korea", country), 
        country=ifelse(country=="Micronesia (Federated States of)", "Micronesia (Fed. States of)", country),
        country=ifelse(country=="Republic of North Macedonia", "North Macedonia", country),
        country=ifelse(country=="West Bank and Gaza Strip", "State of Palestine", country),
        country=ifelse(country=="United Kingdom of Great Britain and Northern Ireland", "United Kingdom", country)
        )
```

```{r dtaPop}
#estimates:Total population, both sexes combined, as of 1 July (thousands)
dtapop<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/WPP2019/WPP2019_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx")

dtapopworld<-dtapop%>%
    rename(country = "Region, subregion, country or area *",
           pop = "2020",
           countrycode = "Country code")%>%
    filter(Type=="World")%>%
    select(country, pop, region)%>%
    mutate(pop=as.numeric(pop))%>%
    mutate(WHO_REGION="")

dtapop<-dtapop%>%
    rename(country = "Region, subregion, country or area *",
           pop = "2020",
           countrycode = "Country code")%>%
    filter(Type=="Country/Area")%>%
    select(country, pop, region)%>%
    mutate(pop=as.numeric(pop))
```

```{r mergePopWHOregion}

dim(dtapop)
dim(dtawhoregion)

temp<-full_join(dtapop, dtawhoregion, by = c("country"))%>%
    mutate(missingpop=is.na(pop))%>%
    filter(missingpop==TRUE)%>%
    arrange(country)

#View(temp) #these are all small territoties and countries

dtapop<-left_join(dtapop, dtawhoregion, by = c("country"))

dim(dtapop)
```

```{r dtaCOVID}
url<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
dtacases<-read.csv(url)%>%
    rename(country = Country.Region,
           region = Province.State)%>%
    gather(variable, value, starts_with("X"))%>%
    rename(date = variable,
           cases = value)%>%
    select(country, date, cases)%>%
    mutate(country=as.character(country))%>%
    group_by(country, date)%>%
    summarize_all(funs(sum))%>%ungroup()

url<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
dtadeaths<-read.csv(url)%>%
    rename(country = Country.Region,
           region = Province.State)%>%
    gather(variable, value, starts_with("X"))%>%
    rename(date = variable,
           deaths = value)%>%
    select(country, date, deaths)%>%
    mutate(country=as.character(country))%>%
    group_by(country, date)%>%
    summarize_all(funs(sum))%>%ungroup()

str(dtacases)
str(dtadeaths)

dtacovid<-left_join(dtacases, dtadeaths, by = c("country", "date"))%>%
    mutate(date=mdy(substring(date, 2)) )

dim(dtacases)
dim(dtadeaths)
dim(dtacovid)

#prep for join/merge
dtacovid<-dtacovid%>%
    mutate(
        country=ifelse(country=="Bolivia","Bolivia (Plurinational State of)", country),
        country=ifelse(country=="Brunei","Brunei Darussalam", country),
        country=ifelse(country=="Burma","Myanmar", country),
        country=ifelse(country=="Congo (Brazzaville)","Congo", country),
        country=ifelse(country=="Congo (Kinshasa)","Democratic Republic of the Congo", country),
        country=ifelse(country=="Cote d'Ivoire","Côte d'Ivoire", country),
        country=ifelse(country=="Iran","Iran (Islamic Republic of)", country),
        country=ifelse(country=="Korea, South","Republic of Korea", country),
        country=ifelse(country=="Laos","Lao People's Democratic Republic", country),
        country=ifelse(country=="Moldova","Republic of Moldova", country),
        country=ifelse(country=="Russia","Russian Federation", country), 
        country=ifelse(country=="Syria","Syrian Arab Republic", country),
        country=ifelse(country=="Taiwan*","China, Taiwan Province of China", country),      
        country=ifelse(country=="Tanzania","United Republic of Tanzania", country),   
        country=ifelse(country=="US","United States of America", country),
        country=ifelse(country=="Venezuela","Venezuela (Bolivarian Republic of)", country),     
        country=ifelse(country=="Vietnam","Viet Nam", country),
        country=ifelse(country=="West Bank and Gaza","State of Palestine", country)
        )
```

```{r dtaCOVIDworld}
url<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
dtacasesworld<-read.csv(url)%>%
    rename(country = Country.Region,
           region = Province.State)%>%
    gather(variable, value, starts_with("X"))%>%
    rename(date = variable,
           cases = value)%>%
    select(date, cases)%>%
    group_by(date)%>%
    summarize_all(funs(sum))%>%ungroup()%>%
    mutate(country="WORLD")

url<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
dtadeathsworld<-read.csv(url)%>%
    rename(country = Country.Region,
           region = Province.State)%>%
    gather(variable, value, starts_with("X"))%>%
    rename(date = variable,
           deaths = value)%>%
    select(date, deaths)%>%
    group_by(date)%>%
    summarize_all(funs(sum))%>%ungroup()%>%
    mutate(country="WORLD")

str(dtacasesworld)
str(dtadeathsworld)

dtacovidworld<-left_join(dtacasesworld, dtadeathsworld, by = c("country", "date"))%>%
    mutate(date=mdy(substring(date, 2)) )

dim(dtacasesworld)
dim(dtadeathsworld)
dim(dtacovidworld)

```

```{r dtaAppendWorld}

str(dtacovidworld)
str(dtacovid)

str(dtapopworld)
str(dtapop)

dtacovid<-rbind(dtacovid, dtacovidworld)
dtapop<-rbind(dtapop, dtapopworld)

dim(dtacovid)
dim(dtapop)

```

```{r dta}
str(dtacovid)
str(dtapop)

dta<-full_join(dtacovid, dtapop, by = "country")

dim(dtacovid)
dim(dtapop)
dim(dta)

temp<-dta%>%filter(is.na(pop)==TRUE) 
table(temp$country) #countries without pop

temp<-dta%>%filter(is.na(cases)==TRUE)
table(temp$country) #countries without COVID data 

dta<-dta%>%
    arrange(country, date)%>%
    filter(is.na(cases)==FALSE & is.na(cases)==FALSE)%>% #countries with POP & COVID data 
    mutate(
        newcases=cases-lag(cases),
        newcases=ifelse(country!=lag(country), NA, newcases),

        newdeaths=deaths-lag(deaths),
        newdeaths=ifelse(country!=lag(country), NA, newdeaths),
        
        incidence=round(100000*cases/(pop*1000), 1), # confiremd cases per 100,000 pop
        cfr=round(100*deaths/cases, 1), # deaths per 100 confirmed cases   
        mortality=round(100000*deaths/(pop*1000), 1) # deaths per 100000 pop
    )%>%
    arrange(country, date)%>%
    mutate(
        newcasessmooth =c(NA,NA,NA,NA,NA,NA,rollmean(newcases, 7)), 
        newdeathssmooth=c(NA,NA,NA,NA,NA,NA,rollmean(newdeaths, 7)), 
        newcasessmooth=ifelse(country !=lag(country ), NA, newcasessmooth),
        newdeathssmooth=ifelse(country !=lag(country ), NA, newdeathssmooth),
        newcasessmooth =round(newcasessmooth, 5),  
        newdeathssmooth=round(newdeathssmooth, 5),

        newcasessmoothpp=round(100*newcasessmooth/pop, 3),   
        newdeathssmoothpp=round(100*newdeathssmooth/pop, 5)
    )

#data quality problem: cumulative deaths and acses going down in some cases 
    
    summary(dta$newcases)
    summary(dta$newdeaths)
    temp<-dta%>%filter(newcases<0 | newdeaths<0)
    dim(temp)
    head(temp, nrow(temp))
    
    temp<-dta%>%filter(newcasessmooth<0 | newdeathssmooth<0)
    dim(temp)
    head(temp, nrow(temp))    

#Force negative smooth numbers to 0 

dta<-dta%>%
    mutate(
        newcasessmooth   =ifelse(newcasessmooth<0,    0, newcasessmooth),
        newcasessmoothpp =ifelse(newcasessmoothpp<0,  0, newcasessmoothpp),
        newdeathssmooth  =ifelse(newdeathssmooth<0,   0, newdeathssmooth),
        newdeathssmoothpp=ifelse(newdeathssmoothpp<0, 0, newdeathssmoothpp)
    )

```

```{r dtacurve}
dtacurve<-dta%>%
    filter(date>="2020-03-01")%>%#only since March
    arrange(country, date)%>%
    group_by(country)%>%
    mutate(
        latest=date==max(date), 
        latestcasespp=newcasessmoothpp, 
        latestcasespp=ifelse(latest==FALSE, NA, latestcasespp)
        )%>%
    fill(latestcasespp)%>%
    fill(latestcasespp, .direction = "up")%>%    
    mutate(
        peakcasespp=round(max(newcasessmoothpp, na.rm = TRUE), 1), 
        peakdate=as.character(date),
            peakdate=ifelse(peakcasespp!=newcasessmoothpp, "", peakdate),
            peakdate=ymd(substring(peakdate, 1)) )%>%
    ungroup()%>%
    select(country, region, WHO_REGION, date, latest, cases, deaths, pop, incidence, cfr, mortality,
           starts_with("new"), starts_with("latest"), starts_with("peak")             )

```

```{r dtaregion}

dtaregion<-dta%>%
    select(country, WHO_REGION, date, cases, deaths, pop)%>% 
    filter(WHO_REGION!="")%>%
    arrange(WHO_REGION, date, country)%>%
    group_by(WHO_REGION, date)%>%
    summarise_at(vars(cases, deaths, pop), funs(sum))%>%
    mutate(
        newcases=cases-lag(cases),
        newcases=ifelse(WHO_REGION!=lag(WHO_REGION), NA, newcases),

        newdeaths=deaths-lag(deaths),
        newdeaths=ifelse(WHO_REGION!=lag(WHO_REGION), NA, newdeaths),
        
        incidence=round(100000*cases/(pop*1000), 1), # confiremd cases per 100,000 pop
        cfr=round(100*deaths/cases, 1), # deaths per 100 confirmed cases   
        mortality=round(100000*deaths/(pop*1000), 1) # deaths per 100000 pop
    )%>%
    arrange(WHO_REGION, date)%>%
    mutate(
        newcasessmooth =c(NA,NA,NA,NA,NA,NA,rollmean(newcases, 7)), 
        newdeathssmooth=c(NA,NA,NA,NA,NA,NA,rollmean(newdeaths, 7)), 
        newcasessmooth=ifelse(WHO_REGION !=lag(WHO_REGION ), NA, newcasessmooth),
        newdeathssmooth=ifelse(WHO_REGION !=lag(WHO_REGION ), NA, newdeathssmooth),
        newcasessmooth =round(newcasessmooth, 5),  
        newdeathssmooth=round(newdeathssmooth, 5),

        newcasessmoothpp=round(100*newcasessmooth/pop, 3),   
        newdeathssmoothpp=round(100*newdeathssmooth/pop, 5)
    )

#data quality problem: cumulative deaths and acses going down in some cases 
    
    summary(dtaregion$newcases)
    summary(dtaregion$newdeaths)
    temp<-dtaregion%>%filter(newcases<0 | newdeaths<0)
    dim(temp)
    head(temp, nrow(temp))
    
    temp<-dtaregion%>%filter(newcasessmooth<0 | newdeathssmooth<0)
    dim(temp)
    head(temp, nrow(temp))    

#Force negative smooth numbers to 0 

dtaregion<-dtaregion%>%
    mutate(
        newcasessmooth   =ifelse(newcasessmooth<0,    0, newcasessmooth),
        newcasessmoothpp =ifelse(newcasessmoothpp<0,  0, newcasessmoothpp),
        newdeathssmooth  =ifelse(newdeathssmooth<0,   0, newdeathssmooth),
        newdeathssmoothpp=ifelse(newdeathssmoothpp<0, 0, newdeathssmoothpp))%>%
    rename_all(function(x) paste0("region", x))%>%
    rename(
        WHO_REGION=regionWHO_REGION, 
        date=regiondate)
    
```

```{r mergeregiondata}
colnames(dtacurve)
colnames(dtaregion)

dim(dtacurve)
dim(dtaregion)
dtacurve<-left_join(dtacurve, dtaregion, by = c("WHO_REGION", "date"))

dim(dtacurve)
colnames(dtacurve)
```

```{r exportcsv}
length(unique(dta$country))
length(unique(dtacurve$country))

write.csv(dtacurve, "dtacurve_global.csv")
#write.csv(dtacurve, "https://github.com/yoonjoung/COVID19_FlattenedCurve/blob/master/dtacurve_global.csv")

dtacurveBrazil<-dtacurve%>%filter(country=="Brazil")
length(unique(dtacurveBrazil$country))
write.csv(dtacurveBrazil, "dtacurve_BrazilExample.csv")
```

###__COVID-19 curves by country__  
(Last updated on `r time`) 

```{r}
ntotalcountries<-(length(unique(dtacurve$country))) - 1 
ntotal<-nrow(dtacurve)
#colnames(dtacurve)
#table(dtacurve$region)
```

###__1. Note on data__

__Source Data__  
1. All COVID-19 data (i.e., cumulative confirmed cases and deaths by day) come from [JHU/CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data). Accessed on `r time`.    
2. All data on country population come from UN [World Population Prospects 2019 Revision](https://population.un.org/wpp/). Accessed on April 18, 2020.   

See note in each source data site about methods and comparability over time and across countries.    

---

__"COVID-19 status data" for dashboard: dtacurve_global.csv__      

* Includes `r ntotalcountries` countries that have __both COVID-19 and population data__.     
* For __global data__, see/filter "country" == "WORLD". This contains global aggregate numbers by date, including all countries/territories that may not be part of the `r ntotalcountries` countries.        
* A total of `r ntotal` country-day-specific observations/rows included, as of `r date`.   

---
   
* Variables/Columns:   
 [1] "country"          
 [2] "region" region classification from UN Pop Division    
 __[3] "WHO_REGION" region classification from WHO* __      
 [4] "date"             
 [5] "latest" latest date reported for the country                    
 [6] "cases" cumulative number of reported cases                         
 [7] "deaths" cumulative number of reported deaths  
 [8] "pop" population in million                 
 [9] "incidence" reported cases per 100,000 pop          
[10] "cfr" deaths per 100 reported cases             
[11] "mortality" deaths per per 100,000 pop        
[12] "newcases"         
[13] "newdeaths"        
[14] "newcasessmooth"   
[15] "newdeathssmooth"  
__[16] "newcasessmoothpp" 7-day rolling average of daily new reported cases per 100,000 pop* __     
__[17] "newdeathssmoothpp" 7-day rolling average of daily new deaths per 100,000 pop* __     
__[18] "latestcasespp" latest 7-day rolling average of daily new reported cases per 100,000 pop* __             
[19] "peakcasespp" highest daily new reported cases per 100,000 pop for the country (based on 7-day rolling average)    
[20] "peakdate"               
<span style="color: #1F77B4;">[21] "regioncases" WHO REGIONAL aggregate            
[22] "regiondeaths" WHO REGIONAL aggregate          
[23] "regionpop" WHO REGIONAL aggregate              
[24] "regionnewcases" WHO REGIONAL aggregate         
[25] "regionnewdeaths" WHO REGIONAL aggregate        
[26] "regionincidence" WHO REGIONAL aggregate       
[27] "regioncfr" WHO REGIONAL aggregate              
[28] "regionmortality" WHO REGIONAL aggregate       
[29] "regionnewcasessmooth" WHO REGIONAL aggregate     
[30] "regionnewdeathssmooth" WHO REGIONAL aggregate    
__[31] "regionnewcasessmoothpp" WHO REGIONAL aggregate* __    
[32] "regionnewdeathssmoothpp" WHO REGIONAL aggregate</span>     

#####Global data 

_Note: The blue and pink lines are new cases and new deaths, respectively. The x-axis is date since March. The LEFT y-axis is 7-day rolling average of daily new cases. The RIGHT y-axis is 7-day rolling average of daily new deaths._  

```{r plot_world, results="asis", fig.align="left", out.width="800px", out.height="400px"}
dtafig<-dtacurve%>%
    filter(country=="WORLD")

mindate<-min(dtacurve$date)
maxdate<-max(dtacurve$date)
maxy<-max(dtafig$newcasessmooth)
maxy2<-max(dtafig$newdeathssmooth)*2

fig1<-dtafig%>% 
    plot_ly(x = ~date)%>% 
    add_lines(
        y = ~newcasessmooth,
        line= list(color = "#395F82"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New cases: ', newcasessmooth)  )%>%
    add_lines(
        y = ~newdeathssmooth,
        yaxis='y2',
        line= list(color = "#9B617F"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New deaths: ', newdeathssmooth)  )%>%    
    add_annotations(
        text = ~unique(country),
        x = 0.5, y = 0.90, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)  )%>%
    layout(
        showlegend = FALSE,
        yaxis=list(title="New cases" , 
                   range=c(0,maxy), showgrid = FALSE,
                   overlaying='y2', side="left"),
        yaxis2 = list(title = "New deaths", 
                      range=c(0,maxy2), showgrid = FALSE,
                      side="right"),
        xaxis=list(title="" , range=c(mindate,maxdate), showgrid = FALSE)
        ) 

maxy<-max(dtafig$cases)
maxy2<-max(dtafig$deaths)*2

fig2<-dtafig%>% 
    plot_ly(x = ~date)%>% 
    add_lines(
        y = ~cases,
        line= list(color = "#395F82"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> Cumulative cases: ', cases)  )%>%
    add_lines(
        y = ~deaths,
        yaxis='y2',
        line= list(color = "#9B617F"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> Cumulative deaths: ', deaths)  )%>%    
    add_annotations(
        text = ~unique(country),
        x = 0.5, y = 0.90, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)  )%>%
    layout(
        showlegend = FALSE,
        yaxis=list(title="Cumulative cases" , 
                   range=c(0,maxy), showgrid = FALSE,
                   overlaying='y2', side="left"),
        yaxis2 = list(title = "Cumulative deaths", 
                      range=c(0,maxy2), showgrid = FALSE,
                      side="right"),
        xaxis=list(title="" , range=c(mindate,maxdate), showgrid = FALSE)
        ) 

fig1

```

#####Country examples

_Note: The blue, pink, and gray lines are new cases, new deaths, and regional new cases, respectively. The x-axis is date since March. The LEFT y-axis is 7-day rolling average of daily new cases per 100,000 population. The RIGHT y-axis is 7-day rolling average of daily new deaths per 100,000 population._  

```{r plot_Brazil, results="asis", fig.align="left", out.width="800px", out.height="400px"}
dtafig<-dtacurve%>%
    mutate(latestcasespp=round(ifelse(latest==FALSE, NA, latestcasespp), 0))%>%
    filter(country=="Brazil")

mindate<-min(dtacurve$date)
maxdate<-max(dtacurve$date)
maxy<-max(dtafig$newcasessmoothpp)
maxy2<-max(dtafig$newdeathssmoothpp)

dtafig%>% 
    plot_ly(x = ~date)%>% 
    add_lines(
        y = ~regionnewcasessmoothpp,
        line= list(color = "#C1C1C1"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Region: ', WHO_REGION,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', regionnewcasessmoothpp)  )%>%
    add_lines(
        y = ~newcasessmoothpp,
        line= list(color = "#395F82"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', newcasessmoothpp)  )%>%
    add_lines(
        y = ~newdeathssmoothpp,
        yaxis='y2',
        line= list(color = "#9B617F"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New deaths per 100,000: ', newdeathssmoothpp)  )%>%    
    add_trace(
        y = ~latestcasespp, type = 'scatter', mode = 'markers',
        marker = list(size = 1,color ="#073763"),
        text = ~latestcasespp, 
        textfont=list(size=10, color="#073763"), 
        textposition = "bottom left"
        )%>%   
    add_annotations(
        text = ~unique(country),
        x = 0.5, y = 0.90, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)  )%>%
    layout(
        showlegend = FALSE,
        yaxis=list(title="New cases per 100,000" , 
                   range=c(0,maxy), showgrid = FALSE,
                   overlaying='y2', side="left"),
        yaxis2 = list(title = "New deaths per 100,000", 
                      range=c(0,2), showgrid = FALSE,
                      side="right"),
        xaxis=list(title="" , range=c(mindate,maxdate), showgrid = FALSE)
        ) 


```

```{r plot_Kenya, results="asis", fig.align="left", out.width="800px", out.height="400px"}
dtafig<-dtacurve%>%
    mutate(latestcasespp=round(ifelse(latest==FALSE, NA, latestcasespp), 0))%>%
    filter(country=="Kenya")

mindate<-min(dtacurve$date)
maxdate<-max(dtacurve$date)
maxy<-max(dtafig$newcasessmoothpp)
maxy2<-max(dtafig$newdeathssmoothpp)

dtafig%>% 
    plot_ly(x = ~date)%>% 
    add_lines(
        y = ~regionnewcasessmoothpp,
        line= list(color = "#C1C1C1"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Region: ', WHO_REGION,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', regionnewcasessmoothpp)  )%>%
    add_lines(
        y = ~newcasessmoothpp,
        line= list(color = "#395F82"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', newcasessmoothpp)  )%>%
    add_lines(
        y = ~newdeathssmoothpp,
        yaxis='y2',
        line= list(color = "#9B617F"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New deaths per 100,000: ', newdeathssmoothpp)  )%>%    
    add_trace(
        y = ~latestcasespp, type = 'scatter', mode = 'markers',
        marker = list(size = 1,color ="#073763"),
        text = ~latestcasespp, 
        textfont=list(size=10, color="#073763"), 
        textposition = "bottom left"
        )%>%   
    add_annotations(
        text = ~unique(country),
        x = 0.5, y = 0.90, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)  )%>%
    layout(
        showlegend = FALSE,
        yaxis=list(title="New cases per 100,000" , 
                   range=c(0,5), showgrid = FALSE,
                   overlaying='y2', side="left"),
        yaxis2 = list(title = "New deaths per 100,000", 
                      range=c(0,1), showgrid = FALSE,
                      side="right"),
        xaxis=list(title="" , range=c(mindate,maxdate), showgrid = FALSE)
        ) 
```

###__2. COVID-19 curve by country in select regions__

```{r plotpanel}
mindate<-min(dtacurve$date)
maxdate<-max(dtacurve$date)

panel <- . %>% 
    plot_ly(x = ~date)%>% 
    add_trace(
        y = 0, type = 'scatter', mode = 'lines',
        line= list(color = "#F1F1F1") ) %>%    
    add_trace(
        y = 10, type = 'scatter', mode = 'lines',
        fill = 'tonexty',fillcolor='#EEEEEE', opacity=0.1,
        line= list(color = "#F1F1F1") ) %>%    
    add_lines(
        y = ~regionnewcasessmoothpp,
        line= list(color = "#C1C1C1"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Region: ', WHO_REGION,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', regionnewcasessmoothpp)  )%>%
    add_lines(
        y = ~newcasessmoothpp,
        line= list(color = "#395F82"),  
        hoverinfo = 'text',
        text = ~paste(
                      '</br> Country: ', country,          
                      '</br> Date: ', date,
                      '</br> New cases per 100,000: ', newcasessmoothpp)  )%>%
    add_trace(
        y = ~latestcasespp, type = 'scatter', mode = 'markers',
        marker = list(size = 1,color ="#073763"),
        text = ~latestcasespp, 
        textfont=list(size=10, color="#073763"), 
        textposition = "bottom left"
        )%>%   
    add_annotations(
        text = ~unique(country),
        x = 0.5, y = 0.90, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)  )%>%
    layout(
        showlegend = FALSE,
        yaxis=list(title="New cases per 100,000" , 
                   range=c(0,maxy), showgrid = FALSE),
        xaxis=list(title="" , range=c(mindate,maxdate), showgrid = FALSE)
        ) 
```
_Note: The x-axis is date since March. The y-axis is 7-day rolling average of daily new cases per 100,000 population. The light gray box represents the number of daily new cases 10 or lower per 100,000._  

####__AFRO__

```{r plot_SouthernAfrica, results="asis", fig.align="left", out.width="800px", out.height="2200px"}

dtafig<-dtacurve%>%
    mutate(latestcasespp=round(ifelse(latest==FALSE, NA, latestcasespp), 0))%>%
    filter(WHO_REGION=="AFRO")

maxy<-max(dtafig$newcasessmoothpp)
nobs<-ceiling(length(unique(dtafig$country))/5)

dtafig%>%
    group_by(country) %>%
    do(p = panel(.)) %>%
    subplot(nrows = nobs, shareX = TRUE, shareY = TRUE)  
```

---

####__PAHO__ 

```{r plot_SouthAmerica, results="asis", fig.align="left", out.width="800px", out.height="1600px"}

dtafig<-dtacurve%>%
    mutate(latestcasespp=round(ifelse(latest==FALSE, NA, latestcasespp), 0))%>%
    filter(WHO_REGION=="PAHO")

maxy<-max(dtafig$newcasessmoothpp)
nobs<-ceiling(length(unique(dtafig$country))/5)

dtafig%>%
    group_by(country) %>%
    do(p = panel(.)) %>%
    subplot(nrows = nobs, shareX = TRUE, shareY = TRUE)  
```

---

<p style="color:gray">
See [GitHub](https://github.com/yoonjoung/COVID19_FlattenedCurve) for data, code, and more information. 
For typos, errors, and questions, contact me at [www.isquared.global](https://www.iSquared.global/YJ). 

_Making Data Delicious, One Byte at a Time_, in good times and bad times.</p>